### 2012-04-26 docs/architecture/architecture.tex
another MAC fix
some updates

\section{\maestro\ Directory Structure}

\begin{figure}[]
\centering
\includegraphics[scale=0.55]{\archfigpath/maestro_ecosystem2}
\caption[\maestro\ directory structure]
{The basic \maestro\ directory structure.}
\end{figure}

Building \maestro\ requires both the \maestro-specific source
files (distributed in the {\tt MAESTRO/} directory), and the
\boxlib\ library (distributed in the {\tt BoxLib/} directory.
\boxlib\ provides both a C++ and a Fortran framework.  \maestro\
only uses the Fortran framework.

@@

@@

\begin{figure}[t]
\centering
\includegraphics[scale=0.65]{\archfigpath/boxlib_directory2}
\caption[\boxlib\ directory structure]
{The basic \boxlib\ directory structure.  The directories used
by \maestro\ are indicated in red.}
\end{figure}

The \boxlib\ directories that \maestro\ uses are:
\begin{itemize}
\item {Src/}

@@

Different \maestro\ problems are defined in sub-directories under
{\tt MAESTRO/}.  To add a problem, start by creating a new
sub-directory---this is where you will compile your problem and
store all the problem-specific files.

The minimum requirement to
define a new problem would be a {\tt GNUmakefile} which describes how
to build the application and an input file which lists the runtime
parameters.  The problem-specific executable is built in the problem
directory by typing {\tt make}.  Source files are found automatically
by searching the directories listed in the {\tt GNUmakefile}.
Customized versions of any source files placed in the problem-directory
override those with the same name found elsewhere.  Any unique
source files (and not simply a custom version of a file found
elsewhere) needs to be listed in a file called {\tt GPackage.mak}
in the problem-directory.

@@

The next line sets the compiler to be used for compilation:
\begin{lstlisting}[language={[gnu]make},mathescape=false]
  COMP := Intel
\end{lstlisting}
The \maestro\ build system knows what options to use for various
compiler families.  The {\tt COMP} flag specifies which compiler to
use.  Allowed values include {\tt Intel}, {\tt gfortran}, {\tt PGI},
and {\tt PathScale}.  The specific details of these choices are
defined in the {\tt MAESTRO/mk/comps/} directory.

@@

Runtime parameters listed in the {\tt MAESTRO/\_parameters} file are
parsed at compile time and the file {\tt probin.f90} is written and
compiled.  This is a Fortran module that holds the values of the
runtime parameters and makes them available to any routine.  The
variable {\tt PROBIN\_PARAMETERS} should list all of the files that
contain runtime parameters for the current problem.  You must always
include the main {\tt ../\_parameters} file (i.e.\ {\tt
  MAESTRO/\_parameters}).  Sometimes a problem defines their own
runtime parameters (for instance {\tt \_parameters.myproblem}), in which
case, its parameter file should also appear in {\tt
  PROBIN\_PARAMETERS} (see \S~\ref{sec:def_runtime_param}).
\begin{lstlisting}[language={[gnu]make},mathescape=false]
  PROBIN_PARAMETERS := ../_parameters
\end{lstlisting}

@@

The runtime parameters for the core \maestro\ algorithm are listed
in {\tt MAESTRO/\_parameters}.  That file is parsed at compile-time
by the {\tt MAESTRO/write\_probin.py} script
(along with any other parameter files listed in the {\tt PROBIN\_PARAMETERS}
variable in the {GNUmakefile}, see \S \ref{sec:makefile}).  The script
outputs the {\tt probin.f90} source file.  Each line
in the {\tt \_parameters} file has the form: 
\vskip 3mm
{\em parameter} \hskip 10em  {\em data-type} \hskip 10em  {\em value} 
\vskip 3mm
\noindent where {\em parameter} is the name of the runtime parameter,
{\em data-type} is one of \{{\tt character}, {\tt real}, {\tt
  integer}, {\tt logical}\}, and the {\em value} specifies the default
value for the runtime parameter.  Comments are indicated by a `{\tt
  \#}' character and are used to produce documentation about the
available runtime parameters.  For the documentation, runtime parameters are grouped together
in the {\tt \_parameters} file into categories.  The category headings
are defined by comments in the {\tt \_parameters} file and any comments
following that heading are placed into that category.  The documentation
is produced by the script {\tt MAESTRO/docs/runtime\_parameters/rp.py}.

@@

Problem-specific runtime parameters should be defined in the
problem-directory in a file called {\tt
  \_parameters}\{{\em.problem-name}\}.  This file should be added
to the {\tt PROBIN\_PARAMETERS} variable in the {\tt GNUmakefile}.

@@

          ! compute cell-centered velocity
          u = HALF*(umac(i,j) + umac(i+1,j))
          v = HALF*(vmac(i,j) + vmac(i+1,j))

>>>>

\section{The \maestro\ `Ecosystem'}

\begin{figure}[]
\centering
\includegraphics[scale=0.55]{\archfigpath/maestro_ecosystem2}
\caption[\maestro\ `ecosystem'] {\label{fig:arch:eco} The basic
  \maestro\ `ecosystem'.  Here we see the different packages that
  contribute to building the {\tt test2} problem in \maestro.  The
  purple lines show the directories that are pulled in through
  variable Makefile variables ({\tt BOXLIB\_HOME}, {\tt NETWORK\_DIR},
  {\tt EOS\_DIR}, and {\tt CONDUCTIVITY\_DIR}).  The blue lines show
  standard modules that are used in most \maestro\ builds.}
\end{figure}

Building \maestro\ requires both the \maestro-specific source
files (distributed in the {\tt MAESTRO/} directory), and the
\boxlib\ library (distributed separately, consisting of the {\tt BoxLib/} directory).
\boxlib\ provides both a C++ and a Fortran framework.  \maestro\
only uses the Fortran portions of \boxlib.  Figure~\ref{fig:arch:eco}
shows the relationship between the different packages, highlighting
what goes into defining a specific \maestro\ problem.

@@

  \item {\tt Source_SDC}

    Source for an alternate formulation of \maestro\ that removes the
    splitting between hydrodynamics, diffusion, and reactions (see
    Chapter~\ref{ch:sdc}).

@@

\begin{figure}[t]
\centering
\includegraphics[scale=0.65]{\archfigpath/boxlib_directory2}
\caption[\boxlib\ directory structure] {\label{fig:arch:boxlib} The
  basic \boxlib\ directory structure.  The directories used by
  \maestro\ are indicated in red.}
\end{figure}

The \boxlib\ directory structure is shown in
Figure~\ref{fig:arch:boxlib}.  The subset of the directories that are
used by \maestro\ are:
\begin{itemize}
\item {Src/}

@@

Different \maestro\ problems are defined in sub-directories under {\tt
  MAESTRO/SCIENCE}, {\tt MAESTRO/TEST\_PROBLEMS}, or {\tt
  MAESTRO/UNIT\_TESTS}.  To add a problem, start by creating a new
sub-directory---this is where you will compile your problem and store
all the problem-specific files.

The minimum requirement to define a new problem would be a {\tt
  GNUmakefile} which describes how to build the application and an
input file which lists the runtime parameters.  The problem-specific
executable is built in the problem directory by typing {\tt make}.
Source files are found automatically by searching the directories
listed in the {\tt GNUmakefile}.  Customized versions of any source
files placed in the problem-directory override those with the same
name found elsewhere.  Any unique source files (and not simply a
custom version of a file found elsewhere) needs to be listed in a file
called {\tt GPackage.mak} in the problem-directory (and this needs to
be told to the build system---see below).

@@

The next line sets the compiler to be used for compilation:
\begin{lstlisting}[language={[gnu]make},mathescape=false]
  COMP := gfortran
\end{lstlisting}
The \maestro\ build system knows what options to use for various
compiler families.  The {\tt COMP} flag specifies which compiler to
use.  Allowed values include {\tt Intel}, {\tt gfortran}, {\tt PGI},
{\tt PathScale}, and {\tt Cray}.  The specific details of these
choices are defined in the {\tt MAESTRO/mk/comps/} directory.

@@

Runtime parameters listed in the {\tt MAESTRO/\_parameters} file are
parsed at compile time and the file {\tt probin.f90} is written and
compiled.  This is a Fortran module that holds the values of the
runtime parameters and makes them available to any routine.  By
default, the build system looks for a file called {\tt \_parameters}
in the problem directory and adds those parameters along with the
master list of \maestro\ parameters ({\tt MAESTRO/\_parameters}) to
the {\tt probin\_module}.

@@

The runtime parameters for the core \maestro\ algorithm are listed in
{\tt MAESTRO/\_parameters}.  That file is parsed at compile-time by
the {\tt MAESTRO/write\_probin.py} script (along with any
problem-specific parameters).  The script outputs the {\tt probin.f90}
source file.  Each line in the {\tt \_parameters} file has the form:
\vskip 3mm
{\em parameter} \hskip 10em  {\em data-type} \hskip 10em  {\em value} 
\vskip 3mm
\noindent where {\em parameter} is the name of the runtime parameter,
{\em data-type} is one of \{{\tt character}, {\tt real}, {\tt
  integer}, {\tt logical}\}, and the {\em value} specifies the default
value for the runtime parameter.  Comments are indicated by a `{\tt
  \#}' character and are used to produce documentation about the
available runtime parameters.  For the documentation, runtime parameters are grouped together
in the {\tt \_parameters} file into categories.  The category headings
are defined by comments in the {\tt \_parameters} file and any comments
following that heading are placed into that category.  The documentation
(Chapter~\ref{ch:parameters}) is produced by the script 
{\tt MAESTRO/docs/runtime\_parameters/rp.py}.

@@

Problem-specific runtime parameters should be defined in the
problem-directory in a file called {\tt \_parameters}.  This file will
be automatically found at compile time.

@@

          ! compute cell-centered velocity
          u = HALF*(umac(i,j) + umac(i+1,j))
          v = HALF*(vmac(i,j) + vmac(i,j+1))

### 2012-04-26 docs/getting_started/getting_started.tex
some updates

\subsection{{\tt data\_processing} scripts}

Several useful analysis scripts (written in Fortran 90) can be found
in {\tt fParallel/data\_processing}.  The {\tt GNUmakefile} there
needs to be edited to indicate which of the tools to build.  For
example, to extract the density along a line from the center of a
plotfile, {\tt plt00200}, in the $y$-direction:

@@

There is also a python visualization method in {\tt
fParallel/data\_processing/python\_plotfile}.  This is described
in \S~\ref{sec:vis:python}.

@@

\visit\ is a powerful, DOE-supported visualization tool for 2- and 3D
datasets.  It can do contouring, volume rendering, streamlines, ...\, ,
directly from \boxlib\ plotfiles (to open a plotfile in \visit, point to
the {\tt Header} file inside the plotfile directory).  Details on
\visit\ can be found at:\newline
 {\tt https://wci.llnl.gov/codes/visit/home.html}\,. \newline
The easiest way to get started with \visit\ is to download a precompiled
binary from the \visit\ webpage.

@@

By default, \maestro\ outputs global diagnostics each timestep into a file
called {\tt maestro\_diag.out}.  This includes the maximum Mach number,
peak temperature, and peak nuclear energy generation rate.  Individual problems
can provide their own {\tt diag.f90} file to produce custom diagnostic
output.

>>>>

\subsection{{\tt AmrPostprocessing} scripts}

Several useful analysis scripts (written in Fortran 90) can be found
in {\tt AmrPostprocessing/F\_Src/} (distributed separately
from \maestro).  The {\tt GNUmakefile} there needs to be edited to
indicate which of the tools to build.  For example, to extract the
density along a line from the center of a plotfile, {\tt plt00200}, in
the $y$-direction:

@@

There is also a python visualization method in {\tt
AmrPostprocessing/F\_src/python\_plotfile}.  This is described
in \S~\ref{sec:vis:python}.

@@

\visit\ is a powerful, DOE-supported visualization tool for 2- and 3D
datasets.  It can do contouring, volume rendering, streamlines, ...\, ,
directly from \boxlib\ plotfiles.   Details on
\visit\ can be found at:\newline
 {\tt https://wci.llnl.gov/codes/visit/home.html}\,. \newline
The easiest way to get started with \visit\ is to download a precompiled
binary from the \visit\ webpage.

@@

By default, \maestro\ outputs global diagnostics each timestep into a
file called {\tt maestro\_diag.out}.  This includes the maximum Mach
number, peak temperature, and peak nuclear energy generation rate.
Individual problems can provide their own {\tt diag.f90} file to
produce custom diagnostic output.  This information can be plotted
directly with {\sl GNUplot}, for example.
